### 2022-07-18 Article.tex
Доработка английского реферата.

	The motion of a near-autonomous time-periodic two-degree-of-freedom Hamil-tonian system is considered in a neighborhood of the trivial equilibrium position being stable in the linear approximation. The system is assumed to implement two Raman third-order resonances being strong and weak. In this case, the frequencies of small oscillations are also related by the fourth-order resonant ratio. In the complete nonlin-ear system, this equilibrium position is unstable. By the perturbation theory methods, normalization of the Hamiltonian of perturbed motion is carried out in the members up to the fourth order inclusive with respect to disturbance, taking into account the existing resonances. The approximate (model) the system depending on three parame-ters is considered under the assumption that the coefficients of terms corresponding to the fourth-order resonance are small. By means of analytical and graphical methods, the issue of the existence and number of equilibrium positions of the model system is obtained. In the space of parameters of the problem, the domains are identified in which the number of equilibrium positions of the model system can vary from zero to five. As an application, the problem of motion of a dynamically symmetric satellite (modeled by a rigid body) about the center of mass in the central Newtonian gravita-tional field on the orbit with small eccentricity is considered. The motion of a satellite is studied in the neighborhood of its periodic motion generating from the hyperboloi-dal precession on a circular orbit. For parameter values corresponding to the multiple Raman third-order resonance, the coefficients of the normal forms of the model sys-tem are found. It is shown that there are three equilibrium points in this system, two of which are unstable and the third stable in the linear approximation. (292)

>>>>

	In this paper, we study the periodic motions of a non-autonomous $2\pi$-time-periodic Hamiltonian system containing a small parameter $\veps$ with two degrees of freedom in a neighborhood of a trivial equilibrium that is stable in the linear approximation. It is assumed that for $\veps = 0$ the Hamiltonian of the system does not contain time, that is, the system is close to autonomous. Let there be no resonances of the first and second orders in the system, and at the same time the values of the parameters are close to the resonance values corresponding to the double (fundamental and combination) resonance of the third order. It is assumed that there is a resonant detuning in one of the frequencies of the linear oscillations of the system. The aim of the paper is to solve the question of the existence of a system of periodic motions in a small neighborhood of an unstable trivial equilibrium, their number and stability. With the help of a number of canonical transformations, the Hamiltonians characteristic of each case close to the resonance one were reduced to a characteristic form. Model systems corresponding to autonomous systems are studied in a sufficiently small neighborhood of the origin. A partition of the parameter space of the problem into regions is found, in each of which the question of the existence and number of resonant periodic motions in a small neighborhood of equilibrium is solved. The results are compared with the case of exact resonance. Using the Sylvester criterion, sufficient conditions for the stability of equilibrium positions of model systems were verified and the question of the number of equilibrium positions satisfying them was solved. The characteristic equation of the linearized system of equations of perturbed motion was analyzed and conditions were found that ensure stability in the linear approximation. At the same time, the question of the number of equilibrium positions of the model system for which one of the conditions is satisfied is solved. A complete analysis of the necessary conditions for stability has not been carried out due to cumbersomeness. Using the Poincare small parameter method, periodic solutions of the complete system are constructed that are analytic in $\veps$ and have a period equal to $12\pi$ in t, and conclusions are drawn about their stability or instability.
